Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-18 sigcse-17/410.tex
sigcse: save some space

This project and the remaining ones are written in
C with a bit of x86-32 assembly code,
which is then
compiled and linked into an ELF executable,
stored into a 1.44-megabyte 3.5-inch floppy-disk image,
and booted via GRUB.
If the image is copied to a real floppy or CD,
%or \shortversion{compact disc,}{embedded into an ``El Torito'' bootable compact disc image,}
it can be booted on standard PC hardware.
However, students most often use
the Simics system emulator \cite{simics} (described below).

@@

\subsection{Simics}

@@

At present the course accepts the following limitations
or core assumptions.

@@

\subsubsection{Automated tests}

%While developing their code,
Students will uncover both bugs
and design misunderstandings by running test code we
provide.
%Some tests are simple, exercising one feature or condition
%(e.g., allocating and freeing a large amount of memory),
%while other tests exercise a mix of kernel features
%simultaneously (creating and destroying processes,
%some of which do I/O while others handle or are killed
%by various exceptions) over an extended period of time.
The test suite is a number of user-space programs
which test
%functionality including
basic thread and process lifecycles, robustness
against corner-case security vulnerabilities, and long-term stability
under large stress tests.
%
%When students submit their thread libraries and kernels,
%we use an automated testing harness to
%run a superset of the public test suite.
%
Of course, due to the nondeterministic nature of many bugs in student code,
these tests can provide only a rough estimate of a submission's correctness,
rather than verifying it with certainty.
(This hit-or-miss nature motivated our work on Landslide.)
%Nevertheless, performance on these tests constitutes 30\% of the kernel project's grade.

\subsubsection{Hurdle completion criteria}

Because of the ambitious learning objectives,
we employ a very detailed notion of completeness
for the thread library and kernel projects.
The ``hurdle'' model is based on the following observations:

% TODO: Figure out how to left-justify these bullet poince.
\begin{itemize}
	\item Students learn more from bugs
		that are harder to find.
		In extreme cases, code may do
		almost everything right, but the last bug may
		demonstrate a deep misunderstanding which requires a re-design.

	\item Because concurrency bugs are triggered rarely,
		finding them requires code capable of running
		for a long period without crashing.

	\item Robust code often requires a different structure
		than a ``usually works'' solution.
		It is easy to %write task-exit code which
		implement \x{vanish()} by
		allocating memory to notify the parent task.
		%parent that it has exited.
		However, such a design suffers from ``terminal irony'':
		a program wishing to exit,
		which would free up memory,
		cannot do so if kernel memory has been exhausted.
\end{itemize}

@@

\subsubsection{Code review and interview}

Finally, both to make up for the tests' incompleteness and to provide personalized human feedback,
each student submission is manually graded by a member of course staff.
The instructors and teaching assistants print each group's submitted code and mark it with red ink, auditing it for style, architectural issues, races, and preemptibility.
Furthermore, after the kernel project, each group meets with their grader for a ``kernel interview'', essentially a debriefing for correcting any misconceptions.
% TODO: and what for p2?
Staff discretionary points constitute 20\% of the kernel project's grade.

>>>>

This project and the remaining ones are written in
C with some x86-32 assembly,
which is then
compiled and linked into an ELF executable,
stored into a %1.44-megabyte 3.5-inch
floppy-disk image,
and booted via GRUB.
If the image is copied to a real floppy or CD,
%or \shortversion{compact disc,}{embedded into an ``El Torito'' bootable compact disc image,}
it can boot on standard PC hardware.
However, students most often use
Simics \cite{simics} (described in \sect{\ref{sec:simics}}).

@@

\subsection{Simics}
\label{sec:simics}

@@

%At present the course accepts the following limitations
%or core assumptions.

@@

\textbf{Automated tests.}
%While developing their code,
Students will uncover both bugs
and design misunderstandings by running test code we
provide.
%Some tests are simple, exercising one feature or condition
%(e.g., allocating and freeing a large amount of memory),
%while other tests exercise a mix of kernel features
%simultaneously (creating and destroying processes,
%some of which do I/O while others handle or are killed
%by various exceptions) over an extended period of time.
The test suite is a number of user-space programs
which test
%functionality including
basic thread and process lifecycles, robustness
against corner-case security vulnerabilities, and long-term stability
under large stress tests.
%
%When students submit their thread libraries and kernels,
%we use an automated testing harness to
%run a superset of the public test suite.
%
Of course, due to the nondeterministic nature of many bugs in student code,
these tests can provide only a rough estimate of a submission's correctness,
rather than verifying it with certainty.
(This hit-or-miss nature motivated our work on Landslide.)
%Nevertheless, performance on these tests constitutes 30\% of the kernel project's grade.

\textbf{Hurdle completion criteria.}
Because of the ambitious learning objectives,
we employ a very detailed notion of completeness
for the thread library and kernel projects.
The ``hurdle'' model is based on the following observations:

% TODO: Figure out how to left-justify these bullet poince.
\begin{itemize}
	\item Students learn more from bugs
		that are harder to find.
		In extreme cases, %code may do
		%almost everything right, but the last bug may
		subtle bugs may
		demonstrate a deep misunderstanding which requires a re-design.

	\item %Because concurrency bugs are triggered rarely,
		Finding rare concurrency bugs requires code capable of running
		for a long period without crashing.

	\item Robust code often requires a different structure
		than a ``usually works'' solution.
		It is easy to %write task-exit code which
		implement \x{vanish()} by
		allocating memory. %to notify the parent task.
		%parent that it has exited.
		However, such a design suffers from ``terminal irony'':
		a program wishing to exit,
		which would free up memory,
		cannot do so if memory is exhausted.
\end{itemize}

@@

\textbf{Code review and interview.}
Finally, both to make up for the tests' incompleteness and to provide personalized human feedback,
each student submission is manually graded by a member of course staff.
The instructors and teaching assistants print each group's submitted code and mark it with red ink, auditing it for style, architectural issues, races, and preemptibility.
Furthermore, after the kernel project, each group meets with their grader for a ``kernel interview'', essentially a debriefing for correcting any misconceptions.
% TODO: and what for p2?
Staff discretionary points constitute 20\% of the kernel project's grade.
